(Deep) Induction for GADTs

### 2021-04-02 induction-paper/GADT-induction.tex
Revised Sections 4 and 5 according to discussion with Patty and Daniel.

%This kind of induction principles, which we will call \emph{structural induction},
Structural induction principles cannot be extended to truly nested types, such as $\mathsf{Bush}$.
%To induce over a truly nested type, we need to use a \emph{deep induction} principle~\cite{jp20}.
Instead, for such data types it is necessary to use a \emph{deep induction} principle~\cite{jp20}.
Such a principle, unlike structural induction, inducts over all of the structured data present,
by traversing not just the outer structure with a predicate $\mathsf{P}$,
but also each data element contained in the data type with a custom predicate $\mathsf{Q}$.
This additional predicate is lifted to predicates on any internal structure containing these data, and the resulting predicates on these internal structures are lifted to predicates on any internal structures containing structures at the previous level, and so on, until the internal structures at all levels of the data type definition, including the top level, have been so processed.
Satisfaction of a predicate by the data at one level of a structure is then conditioned upon satisfaction of the appropriate predicates by all of the data at the preceding level.

@@

As we have seen in Subsection~\ref{ssec:induction-nested},
truly nested types do not support a structural induction rule,
which is the reason why it was necessary to introduce a deep induction rule supporting them.
Consequently, GADTs do not support a structural induction rule either,
as they generalize nested types.
Still, there is hope for GADTs to support a deep induction rule,
like nested types do.

@@

\subsection{Induction for GADTs with nesting}\label{sec:GADT-nested}

>>>>

%This kind of induction principles, which we will call \emph{structural induction},
Structural induction principles cannot be extended to truly nested types, such as $\mathsf{Bush}$.
{\color{red} THIS IS NOT TRUE!!!}
%To induce over a truly nested type, we need to use a \emph{deep induction} principle~\cite{jp20}.
Instead, for such data types it is necessary to use a \emph{deep induction} principle~\cite{jp20}.
Such a principle, unlike structural induction, inducts over all of the structured data present,
by traversing not just the outer structure with a predicate $\mathsf{P}$,
but also each data element contained in the data type with a custom predicate $\mathsf{Q}$.
This additional predicate is lifted to predicates on any internal structure containing these data, and the resulting predicates on these internal structures are lifted to predicates on any internal structures containing structures at the previous level, and so on, until the internal structures at all levels of the data type definition, including the top level, have been so processed.
Satisfaction of a predicate by the data at one level of a structure is then conditioned upon satisfaction of the appropriate predicates by all of the data at the preceding level.

@@

As we have seen in Subsection~\ref{ssec:induction-nested},
truly nested types do not support a structural induction rule,
which is the reason why it was necessary to introduce a deep induction rule supporting them.
Consequently, GADTs do not support a structural induction rule either,
as they generalize nested types.
{\color{red} THIS IS NOT TRUE!!!}
Still, there is hope for GADTs to support a deep induction rule,
like nested types do.

@@

Finally, we generalize the approach taken in the previous examples
and provide a general framework to derive induction rules for GADTs.
%Since, as seen in Section~\ref{sec:GADT-nested},
%GADTs featuring nesting do not admit an induction rule,
%we restrict our scope to GADTs not featuring nesting.
For that, we need to give a grammar for the types we will be considering.
A generic GADT
\begin{equation*}
\begin{array}{l}
\mathsf{data\, G\,(\ol{a : Set}) : Set\,where}\\
\mathsf{\;\;\;\;\;\;\;\;C_i\, :\, F_i\,G\,\ol{b} \to G (\ol{K_i\,\ol{b}})}
\end{array}
\end{equation*}
is defined by a finite number of constructors $\mathsf{C_i}$.
In the definition above, $\mathsf{F_i}$ is a type constructor with signature $\mathsf{(Set^{\alpha} \to Set) \to Set^{\beta} \to Set}$
and each $\mathsf{K_i}$ is a type constructor with signature $\mathsf{Set^{\beta} \to Set}$
(i.e. a type constructor of arity $\mathsf{\beta}$).
The overline notation denotes a finite list:
$\mathsf{\ol{a}}$ is a list of types of length $\mathsf{\alpha}$,
so that it can be applied to the type constructor $\mathsf{G}$ of arity $\mathsf{\alpha}$.
%and the vertical bars notation denotes the length of a list:
%$\mathsf{G}$ depends on a finite list of types $\mathsf{\ol{a}}$ of length $\mathsf{\alpha}$,
Each of the $\mathsf{\alpha}$-many $\mathsf{K_i}$ is a type constructor of arity $\mathsf{\beta}$
so that it can be applied to the list of types $\mathsf{\ol{b}}$ of length $\mathsf{\beta}$.
%depends on a finite list of types $\mathsf{\ol{b}}$ of length $\mathsf{\beta}$,
Moreover, notice that the arity of $\mathsf{G}$ matches the number of type constructors $\mathsf{\ol{K_i}}$.
We allow each $\mathsf{F_i}$ to be inductively built in the following ways (and with the following restrictions):
\item $\mathsf{F_i = F_i' \times F_i''}$ where $\mathsf{F_i'}$ and $\mathsf{F_i''}$ have the same signature as $\mathsf{F_i}$ and are built recursively from the same induction rules.
\item $\mathsf{F_i = F_i' + F_i''}$ where $\mathsf{F_i'}$ and $\mathsf{F_i''}$ have the same signature as $\mathsf{F_i}$ and are built recursively from the same induction rules.
\item $\mathsf{F_i = F_i' \to F_i''}$
where $\mathsf{F_i'}$ does not contain the recursive variable,
i.e., $\mathsf{F_i' : Set^\beta \to Set}$,
and $\mathsf{F_i''}$ has the same signature as $\mathsf{F_i}$ and is built recursively from the same induction rules.
\item $\mathsf{F_i\,G\,\ol{b} = G(\ol{F_a\,\ol{b}})}$
Such restriction is necessary to prevent nesting,
as that would break the induction rule as discussed in Section~\ref{sec:GADT-nested}.
\item $\mathsf{F_i\,G\,\ol{b} = H\,\ol{b}}$
Notice that this covers the case in which $\mathsf{F_i}$ is a closed type,
so, in particular, the unit and empty types, $\mathsf{1}$ and $\mathsf{0}$.
%The only requirement on $\mathsf{H}$ is that it needs to have a predicate-lifting function
%(and any type made of constant types, sums, products, arrow types and other GADTs satisfies this condition).
\item $\mathsf{F_i\,G\,\ol{b} = H (\ol{F_c\,G\,\ol{b}})}$
where $\mathsf{H}$ is a $\mathsf{\gamma}$-ary type constructor not containing the recursive variable,
and $\mathsf{F_c}$ has the same signature as $\mathsf{F_i}$  and is built recursively from the same induction rules, for every $\mathsf{c = 1 \dots \gamma}$.
Moreover, we require that $\mathsf{H}$ is not a GADT itself (but we allow it to be an ADT or even a nested type).
We can summarize the above inductive definition with the following grammar
(but beware that the above restrictions and requirements still apply):
\mathsf{F_i\,G\,\ol{b} :=
F_i'\,G\,\ol{b} \times F_i''\,G\,\ol{b}
\ \vert\ F_i'\,G\,\ol{b} + F_i''\,G\,\ol{b}
\ \vert\ F_i'\,\ol{b} \to F_i''\,G\,\ol{b}
\ \vert\ G(\ol{F_a\,\ol{b}})
\ \vert\ H\,\ol{b}
\ \vert\ H (\ol{F_c\,G\,\ol{b}})
}
%The only requirement on the $\mathsf{K_i}$ is that they need to have a predicate-lifting function.

Consider such a generic GADT
\begin{equation}\label{eq:gadt}
\begin{array}{l}
\mathsf{data\, G\,(a : Set) : Set\,where}\\
\mathsf{\;\;\;\;\;\;\;\;C\, :\, F\,G\,\ol{b} \to G (\ol{K\,\ol{b}})}
\end{array}
\end{equation}
which, for ease of notation,
we assume to be a unary type constructor (i.e. it depends on a single type parameter $\mathsf{a}$)
and to have only one constructor $\mathsf{C}$.
Extending the argument to GADTs of arbitrary arity and with multiple constructors presents no difficulty.
In the definition above, $\mathsf{F}$ has type $\mathsf{(Set \to Set) \to Set^{\beta} \to Set}$
and each $\mathsf{K}$ has type $\mathsf{Set^{\beta} \to Set}$.
%The overline notation denotes a finite list:
%%and the vertical bars notation denotes the length of a list:
%%$\mathsf{G}$ depends on a finite list of types $\mathsf{\ol{a}}$ of length $\mathsf{\alpha}$,
%each of the $\mathsf{\alpha}$-many $\mathsf{K}$ depends on a finite list of types $\mathsf{\ol{b}}$ of length $\mathsf{\beta}$,
%and notice that the length of the list of types $\mathsf{\ol{a}}$ matches that of the list of type constructors $\mathsf{\ol{K}}$.
The constructor $\mathsf{C}$ can be rewritten as
\[
\mathsf{C : \exists (\ol{b : Set}) \to Equal\,a\,(K\,\ol{b}) \to F\,G\,\ol{b} \to G\,a}
\]
which is the form we shall use from now on.
%Notice that, in the above constructor, we have a type
%$\mathsf{Equal\,a\,(K\,\ol{b})}$ for each $\mathsf{a}$ and its corresponding type constructor $\mathsf{K}$.

In order to state the induction rule for $\mathsf{G}$,
we first need to define $\mathsf{G}$'s associated predicate-lifting function 
\[
\mathsf{G^{\wedge} : \forall (a : Set) \to (a \to Set) \to G\,a \to Set}
\]
as
\[
\mathsf{G^{\wedge}\,a\,Q_a\,(C\,\ol{b}\,e\,x)
= \exists (\ol{Q_b : b \to Set})
\to Equal^{\wedge}\,a\,(K\,\ol{b})\,Q_a\,(K^{\wedge}\,\ol{b}\,\ol{Q_b})\,e
\times F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,{\ol{Q_b}}\,x}
\]
where $\mathsf{Q_a : a \to Set}$
and $\mathsf{C\,\ol{b}\,e\,x : G\,a}$,
i.e., $\mathsf{e : Equal\,a\,(K\,\ol{b})}$ and $\mathsf{x : F\,G\,\ol{b}}$.
%Notice that we need a type $\mathsf{Equal^{\wedge}\,a\,(K\,\ol{b})\,Q_a\,(K^{\wedge}\,\ol{b}\,\ol{Q_b})\,e_a}$
%for each type $\mathsf{a}$ and its corresponding type constructor $\mathsf{K}$.
We also assume to have liftings for $\mathsf{F}$,
\begin{multline*}
\mathsf{F^{\wedge} : \forall (G : Set^{\alpha} \to Set) (\ol{b : Set})
\to (\forall (a : Set) \to (a \to Set) \to G\,a \to Set)} \\
\mathsf{\to (\ol{b \to Set})
\to F\,G\,\ol{b} \to Set}
\end{multline*}
and for  $\mathsf{K}$,
\[
\mathsf{K^{\wedge} : \forall (\ol{b : Set}) \to (\ol{b \to Set}) \to K\,\ol{b} \to Set}
\]
as we could not possibly define an induction rule otherwise.
A way to concretely define the predicate-lifting function for a type is to proceed by induction on the structure of the type,
and we have seen in the previous sections examples of how to do so for products and type application.
We do not give here the general definition of lifting,
as that would require to first present a full type calculus,
and that is beyond the scope of the paper.

Finally, associate the type
\begin{multline*}
\mathsf{CC = \forall (P : \forall (a : Set) \to (a \to Set) \to G\,a \to Set)} \\
\mathsf{\to \forall (a : Set) (\ol{b : Set}) (Q_a : a \to Set) (\ol{Q_b : b \to Set}) (e : Equal\,a\,(K\,\ol{b})) (x : F\,G\,\ol{b})} \\
\mathsf{\to Equal^{\wedge}\,a\,(K\,\ol{b})\,Q_a\,(K^{\wedge}\,\ol{b}\,\ol{Q_b})\,e
	\to F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x
	\to P\,a\,Q_a\,(C\,\ol{b}\,e\,x)}
\end{multline*}
to the constructor $\mathsf{C}$.
\mathsf{DIG\,P\,cc\,a\,Q_a\,(C\,\ol{b}\,e\,x)\,(\ol{Q_b}, L_E, L_F)
= cc\,a\,\ol{b}\,Q_a\,\ol{Q_b}\,e\,x\,L_E\,(p\,x\,L_F)}
\item $\mathsf{C\,\ol{b}\,e\,x : G\,a}$,
i.e., $\mathsf{e : Equal\,a\,(K\,\ol{b})}$,
\item $\mathsf{(\ol{Q_b}, L_E, L_F) : G^{\wedge}\,a\,Q_a (C\,\ol{b}\,e\,x)}$,
$\mathsf{L_E : Equal^{\wedge}\,a\,(K\,\ol{b})\,Q_a\,(K^{\wedge}\,\ol{b}\,\ol{Q_b})\,e}$,
\item Case $\mathsf{F = F_1 \times F_2}$ where $\mathsf{F_1}$ and $\mathsf{F_2}$ have the same signature as $\mathsf{F}$.
\item Case $\mathsf{F = F_1 + F_2}$ where $\mathsf{F_1}$ and $\mathsf{F_2}$ have the same signature as $\mathsf{F}$.
%where $\mathsf{F_1}$ and $\mathsf{F_2}$ have the same type as $\mathsf{F}$.
where $\mathsf{F_1}$ does not contain the recursive variable,
i.e., $\mathsf{F_1 : Set^\beta \to Set}$,
and $\mathsf{F_2}$ has the same signature as $\mathsf{F}$.
\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x = \forall (z : F_1\,\ol{b}) \to F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,z \to F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,(x\,z)}
%\mathsf{F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x = \forall (z : F_1\,G\,\ol{b}) \to F_2^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,(x\,y)} \\
%%F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,x \to F_2^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x} \\
%\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x = \forall (z : F_1\,G\,\ol{b}) \to F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,(x\,y)}
%%\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x = F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,x \to F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x}
where $\mathsf{x : F\,G\,\ol{b} = F_1\,\ol{b} \to F_2\,G\,\ol{b}}$.
Thus, we define $\mathsf{p\,x\,L_F : F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x}$
for $\mathsf{L_F : F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x}$
as $\mathsf{p\,x\,L_F\,z\,L_1 = p_2 (x\,z) (L_F\,z\,L_1)}$
for $\mathsf{z : F_1\,\ol{b}}$
and $\mathsf{L_1 : F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,z}$.
Notice that $\mathsf{F_1}$ not containing the recursive variable
is a necessary restriction, as the proof relies on
$\mathsf{F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x}$
and $\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x}$
having the same domain.
\item Case $\mathsf{F\,G\,\ol{b} = G(F'\,\ol{b})}$
where $\mathsf{F'}$ does not contain the recursive variable,
i.e., $\mathsf{F' : Set^\beta \to Set}$.
Thus, $\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b} = P (F'\,\ol{b}) (F'^{\wedge}\,\ol{b}\,\ol{Q_b})}$.
\mathsf{p = DIG\,P\,cc\,(F'\,\ol{b})\,(F'^{\wedge}\,\ol{b}\,\ol{Q_b})}
%Note that the fact that the type constructor $\mathsf{F'}$ does not contain the recursive variable
%means that $\mathsf{G}$ does not feature nesting.
%As explained and exemplified in Section~\ref{sec:GADT-nested},
%that assumption is required to make induction work.
where $\mathsf{H}$ is a $\mathsf{\beta}$-ary type constructor not containing the recursive variable,
where $\mathsf{H}$ is a $\mathsf{\gamma}$-ary type constructor not containing the recursive variable,
and $\mathsf{F_c}$ has the same signature as $\mathsf{F}$, for every $\mathsf{c = 1 \dots \gamma}$.
Notice that, by proceeding by structural induction as above,
we have implicitly defined a grammar for GADTs admitting an induction rule.
In particular, we allow $\mathsf{F}$ to feature sums, products, arrow types and (limited) type application.

%\subsection{General case (Daniel's revision)}
%%We restrict our scope to GADTs not featuring nesting,
%%in the sense that their constructors contain no nested calls of the recursive variable,
%%as truly nested types do.
%%This way we can avoid the issue described in Section~\ref{sec:GADT-nested}.
%%The reason for this restriction will be discussed in Section~\ref{sec:GADT-nested}.
%Finally, we generalize the approach taken in the previous examples
%and provide a general framework to derive induction rules for GADTs.
%%Since, as seen in Section~\ref{sec:GADT-nested},
%%GADTs featuring nesting do not admit an induction rule,
%%we restrict our scope to GADTs not featuring nesting.
%For that, we need to give a grammar for the types we will be considering.
%A generic GADT
%\begin{equation}\label{eq:gadt}
%\begin{array}{l}
%\mathsf{data\, G\,(\ol{a : Set}) : Set\,where}\\
%\mathsf{\;\;\;\;\;\;\;\;C_i\, :\, F_i\,G\,\ol{b} \to G (\ol{K_i\,\ol{b}})}
%\end{array}
%\end{equation}
%is defined by a finite number of constructors $\mathsf{C_i}$.
%In the definition above, $\mathsf{F_i}$ has type $\mathsf{(Set^{\alpha} \to Set) \to Set^{\beta} \to Set}$
%and each $\mathsf{K_i}$ has type $\mathsf{Set^{\beta} \to Set}$.
%The overline notation denotes a finite list:
%%and the vertical bars notation denotes the length of a list:
%$\mathsf{G}$ depends on a finite list of types $\mathsf{\ol{a}}$ of length $\mathsf{\alpha}$,
%each of the $\mathsf{\alpha}$-many $\mathsf{K_i}$ depends on a finite list of types $\mathsf{\ol{b}}$ of length $\mathsf{\beta}$,
%and notice that the length of the list of types $\mathsf{\ol{a}}$ matches that of the list of type constructors $\mathsf{\ol{K_i}}$.
%We allow each $\mathsf{F_i}$ to be inductively built in the following ways (and with the following restrictions):
%\begin{itemize}
%\item $\mathsf{F_i = F_i' \times F_i''}$ where $\mathsf{F_i'}$ and $\mathsf{F_i''}$ have the same type as $\mathsf{F_i}$.
%\item $\mathsf{F_i = F_i' + F_i''}$ where $\mathsf{F_i'}$ and $\mathsf{F_i''}$ have the same type as $\mathsf{F_i}$.
%\item $\mathsf{F_i = F_i' \to F_i''}$
%where $\mathsf{F_i'}$ does not contain the recursive variable,
%i.e., $\mathsf{F_i' : Set^\beta \to Set}$,
%and $\mathsf{F_i''}$ has the same type as $\mathsf{F_i}$.
%\item $\mathsf{F_i\,G\,\ol{b} = G(\ol{F_a\,\ol{b}})}$
%where none of the $\mathsf{F_a}$ contains the recursive variable,
%i.e., $\mathsf{F_a : Set^\beta \to Set}$
%for each $\mathsf{a}$.
%Such restriction is necessary to prevent nesting,
%as that would break the induction rule as discussed in Section~\ref{sec:GADT-nested}.
%\item $\mathsf{F_i\,G\,\ol{b} = H\,\ol{b}}$
%where $\mathsf{H}$ is a type constructor not containing the recursive variable,
%i.e., $\mathsf{H : Set^\beta \to Set}$.
%Notice that this covers the case in which $\mathsf{F_i}$ is a closed type,
%so, in particular, the unit and empty types, $\mathsf{1}$ and $\mathsf{0}$.
%The only requirement on $\mathsf{H}$ is that it needs to have a predicate-lifting function
%(and any type made of constant types, sums, products, arrow types and other GADTs satisfies this condition).
%\item $\mathsf{F_i\,G\,\ol{b} = H (\ol{F_c\,G\,\ol{b}})}$
%where $\mathsf{H}$ is a type constructor not containing the recursive variable,
%i.e., $\mathsf{H : Set^\gamma \to Set}$,
%and $\mathsf{F_c}$ has the same type as $\mathsf{F_i}$, for every $\mathsf{c = 1 \dots \gamma}$.
%Moreover, we assume that $\mathsf{H}$ is not a GADT itself (but we allow it to be an ADT or even a nested type).
%\end{itemize}
%The only requirement on the $\mathsf{K_i}$ is that they need to have a predicate-lifting function.
%Consider such a generic GADT
%\begin{equation}\label{eq:gadt}
%\begin{array}{l}
%\mathsf{data\, G\,(\ol{a : Set}) : Set\,where}\\
%\mathsf{\;\;\;\;\;\;\;\;C\, :\, F\,G\,\ol{b} \to G (\ol{K\,\ol{b}})}
%\end{array}
%\end{equation}
%which, for ease of notation,
%we assume to have only one constructor $\mathsf{C}$.
%Extending the argument to GADTs with multiple constructors presents no difficulty.
%In the definition above, $\mathsf{F}$ has type $\mathsf{(Set^{\alpha} \to Set) \to Set^{\beta} \to Set}$
%and each $\mathsf{K}$ has type $\mathsf{Set^{\beta} \to Set}$.
%The overline notation denotes a finite list:
%%and the vertical bars notation denotes the length of a list:
%$\mathsf{G}$ depends on a finite list of types $\mathsf{\ol{a}}$ of length $\mathsf{\alpha}$,
%each of the $\mathsf{\alpha}$-many $\mathsf{K}$ depends on a finite list of types $\mathsf{\ol{b}}$ of length $\mathsf{\beta}$,
%and notice that the length of the list of types $\mathsf{\ol{a}}$ matches that of the list of type constructors $\mathsf{\ol{K}}$.
%With that in mind, the constructor $\mathsf{C}$ can be rewritten as
%\mathsf{C : \exists (\ol{b : Set}) \to \ol{Equal\,a\,(K\,\ol{b})} \to F\,G\,\ol{b} \to G\,\ol{a}}
%which is the form we shall use from now on.
%Notice that, in the above constructor, we have a type
%$\mathsf{Equal\,a\,(K\,\ol{b})}$ for each $\mathsf{a}$ and its corresponding type constructor $\mathsf{K}$.
%In order to state the induction rule for $\mathsf{G}$,
%we first need to define $\mathsf{G}$'s associated predicate-lifting function 
%\mathsf{G^{\wedge} : \forall (\ol{a : Set}) \to (\ol{a \to Set}) \to G\,\ol{a} \to Set}
%\mathsf{G^{\wedge}\,\ol{a}\,\ol{Q_a}\,(C\,\ol{b}\,\ol{e_a}\,x)
%= \exists (\ol{Q_b : b \to Set})
%\to F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,{\ol{Q_b}}\,x}
%where $\mathsf{Q_a : a \to Set}$
%and $\mathsf{C\,\ol{b}\,\ol{e_a}\,x : G\,\ol{a}}$,
%i.e., $\mathsf{e_a : Equal\,a\,(K\,\ol{b})}$ and $\mathsf{x : F\,G\,\ol{b}}$.
%We also assume to have liftings for $\mathsf{F}$,
%\mathsf{F^{\wedge} : \forall (G : Set^{\alpha} \to Set) (\ol{b : Set})
%\to (\forall (\ol{a : Set}) \to (\ol{a \to Set}) \to G\,\ol{a} \to Set)} \\
%\mathsf{\to (\ol{b \to Set})
%\to F\,G\,\ol{b} \to Set}
%and for each  $\mathsf{K}$,
%\mathsf{K^{\wedge} : \forall (\ol{b : Set}) \to (\ol{b \to Set}) \to K\,\ol{b} \to Set}
%as we could not possibly define an induction rule otherwise.
%A way to concretely define the predicate-lifting function for a type is to proceed by induction on the structure of the type,
%and we have seen in the previous sections examples of how to do so for products and type application.
%We do not give here the general definition of lifting,
%as that would require to first present a full type calculus,
%and that is beyond the scope of the paper.
%Finally, associate the type
%\mathsf{CC = \forall (P : \forall (\ol{a : Set}) \to (\ol{a \to Set}) \to G\,\ol{a} \to Set)} \\
%\mathsf{\to \forall (\ol{a : Set}) (\ol{b : Set}) (\ol{Q_a : a \to Set}) (\ol{Q_b : b \to Set}) (\ol{e_a : Equal\,a\,(K\,\ol{b})}) (x : F\,G\,\ol{b})} \\
%\mathsf{\to F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x
%	\to P\,\ol{a}\,\ol{Q_a}\,(C\,\ol{b}\,\ol{e_a}\,x)}
%to the constructor $\mathsf{C}$.
%%We can now define the induction rule for $\mathsf{G}$
%The induction rule for $\mathsf{G}$ is
%\mathsf{\forall (P : \forall (\ol{a : Set}) \to (\ol{a \to Set}) \to G\,\ol{a} \to Set) \to CC\,P} \\
%%\mathsf{\to \Big(
%%\forall (\ol{a : Set}) (\ol{b : Set}) (\ol{Q_a : a \to Set}) (\ol{Q_b : b \to Set}) (\ol{e_a : Equal\,a\,(K\,\ol{b})}) (x : F\,G\,\ol{b})} \\
%%\mathsf{\to \ol{Equal^{\wedge}\,a\,(K\,\ol{b})\,Q_a\,(K^{\wedge}\,\ol{b}\,\ol{Q_b})\,e_a}
%%	\to F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x
%%	\to P\,\ol{a}\,\ol{Q_a}\,(C\,\ol{b}\,\ol{e_a}\,x)
%%\Big)} \\
%\mathsf{\to \forall (\ol{a : Set})(\ol{Q_a : a \to Set})(y : G\,\ol{a}) \to G^{\wedge}\,\ol{a}\,\ol{Q_a}\,y \to P\,\ol{a}\,\ol{Q_a}\,y}
%As always, we now define a term $\mathsf{DIG}$ for the induction rule.
%\mathsf{DIG\,P\,cc\,\ol{a}\,\ol{Q_a}\,(C\,\ol{b}\,\ol{e_a}\,x)\,(\ol{Q_b}, L_F)
%= cc\,\ol{a}\,\ol{b}\,\ol{Q_a}\,\ol{Q_b}\,\ol{e_a}\,x\,(p\,x\,L_F)}
%\begin{itemize}
%\item $\mathsf{cc : CC\,P}$;
%\item $\mathsf{C\,\ol{b}\,\ol{e_a}\,x : G\,\ol{a}}$,
%i.e., $\mathsf{e_a : Equal\,a\,(K\,\ol{b})}$ for each $\mathsf{a}$,
%and $\mathsf{x : F\,G\,\ol{b}}$;
%\item $\mathsf{(\ol{Q_b}, L_F) : G^{\wedge}\,\ol{a}\,\ol{Q_a} (C\,\ol{b}\,\ol{e_a}\,x)}$,
%i.e.,
%$\mathsf{Q_b : b \to Set}$ for each $\mathsf{b}$,
%and $\mathsf{L_F : F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,{\ol{Q_b}}\,x}$.
%\end{itemize}
%Finally, the morphism of predicates
%\mathsf{p : PredMap\,(F\,G\,\ol{b})\,(F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}) (F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b})}
%is defined by structural induction on $\mathsf{F}$ as follows:
%\begin{itemize}
%\item Case $\mathsf{F = F_1 \times F_2}$ where $\mathsf{F_1}$ and $\mathsf{F_2}$ have the same type as $\mathsf{F}$.
%We have that
%\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}
%= Pair^{\wedge}(F_1\,G\,\ol{b})(F_2\,G\,\ol{b})(F_1^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}) (F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b})}
%By inductive hypothesis, there exist morphisms of predicates
%\mathsf{p_1 : PredMap\,(F_1\,G\,\ol{b})\,((F_1)^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}) ((F_1)^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b})} \\
%\mathsf{p_2 : PredMap\,(F_2\,G\,\ol{b})\,((F_2)^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}) ((F_2)^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b})}
%Thus, we define $\mathsf{p (x_1, x_2) (L_1, L_2) = (p_1\,x_1\,L_1, p_2\,x_2\,L_2)}$ 
%for 
%$\mathsf{x_1 : F_1\,G\,\ol{b}}$,
%$\mathsf{x_2 : F_2\,G\,\ol{b}}$,
%$\mathsf{L_1 : F_1^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x_1}$
%and $\mathsf{L_2 : F_2^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x_2}$.
%%%%
%\item Case $\mathsf{F = F_1 + F_2}$ where $\mathsf{F_1}$ and $\mathsf{F_2}$ have the same type as $\mathsf{F}$.
%Analogous to case $\mathsf{F = F_1 \times F_2}$.
%%%%
%\item Case $\mathsf{F = F_1 \to F_2}$
%%where $\mathsf{F_1}$ and $\mathsf{F_2}$ have the same type as $\mathsf{F}$.
%i.e., $\mathsf{F_1 : Set^\beta \to Set}$,
%We have that
%\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x = \forall (z : F_1\,\ol{b}) \to F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,z \to F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,(x\,z)}
%%\begin{gather*}
%%\mathsf{F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x = \forall (z : F_1\,G\,\ol{b}) \to F_2^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,(x\,y)} \\
%%%F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,x \to F_2^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x} \\
%%\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x = \forall (z : F_1\,G\,\ol{b}) \to F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,(x\,y)}
%%%\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x = F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,x \to F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x}
%%\end{gather*}
%where $\mathsf{x : F\,G\,\ol{b} = F_1\,\ol{b} \to F_2\,G\,\ol{b}}$.
%By inductive hypothesis, there exist a morphism of predicates
%\mathsf{p_2 : PredMap\,(F_2\,G\,\ol{b}) (F_2^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}) (F_2^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b})}
%Thus, we define $\mathsf{p\,x\,L_F : F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x}$
%for $\mathsf{L_F : F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x}$
%as $\mathsf{p\,x\,L_F\,z\,L_1 = p_2 (x\,z) (L_F\,z\,L_1)}$
%for $\mathsf{z : F_1\,\ol{b}}$
%and $\mathsf{L_1 : F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,z}$.
%%{\color{blue} Revise this case!}
%%Thus, we define $\mathsf{p\,x\,L_F\,L_1 = p_2\,(L_F\,L_1)}$
%%for $\mathsf{L_1 : F_1^{\wedge}\,\ol{b}\,\ol{Q_b}\,x}$.
%Notice that $\mathsf{F_1}$ does not containing the recursive variable
%is a necessary restriction, as the proof relies on
%%%%
%\item Case $\mathsf{F\,G\,\ol{b} = G(\ol{F_a\,\ol{b}})}$
%where none of the $\mathsf{F_a}$ contains the recursive variable,
%i.e., $\mathsf{F_a : Set^\beta \to Set}$
%for each $\mathsf{a}$.
%Thus, $\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b} = P (\ol{F_a\,\ol{b}}) (\ol{F_a^{\wedge}\,\ol{b}\,\ol{Q_b}})}$.
%So, $\mathsf{p}$ is defined as
%\mathsf{p = DIG\,P\,cc\,(\ol{F_a\,\ol{b}})\,(\ol{F_a^{\wedge}\,\ol{b}\,\ol{Q_b}})}
%%for $\mathsf{L_F : F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x}$.
%Note that the fact that the type constructors $\mathsf{F_a}$ do not contain the recursive variable
%means that $\mathsf{G}$ does not feature nesting.
%As explained and exemplified in Section~\ref{sec:GADT-nested},
%that assumption is required to make induction work.
%%%%
%\item Case $\mathsf{F\,G\,\ol{b} = H\,\ol{b}}$
%where $\mathsf{H}$ is a type constructor not containing the recursive variable,
%i.e., $\mathsf{H : Set^\beta \to Set}$.
%In such case, $\mathsf{p : PredMap (H\,\ol{b}) (H^{\wedge}\,\ol{b}\,\ol{Q_b}) (H^{\wedge}\,\ol{b}\,\ol{Q_b})}$
%is just the identity morphism of predicates.
%%%%
%\item Case $\mathsf{F\,G\,\ol{b} = H (\ol{F_c\,G\,\ol{b}})}$
%where $\mathsf{H}$ is a type constructor not containing the recursive variable,
%i.e., $\mathsf{H : Set^\gamma \to Set}$,
